Meta-analytic approaches and 'new' effect sizes to account for heterogeneous effects in comparative physiology

### 2021-06-14 MS/ms.Rmd
Adding some written stuff. Update dto ref

Despite the widespread adoption and well established methodological procedures of meta-analysis it is often criticised for mixing 'apples and oranges' - in otherwords, mixing effects from studies that lack comparability [@Gurevitch2018; @Arnqvist1995; @Gallo1979; @Carpenter2020; @Stewart2010]. Lack of comparability could result from studies differing in experimental design, species, measurement variables and sampling units, which can impact the generalisability of meta-analytic results [@Arnqvist1995; @Stewart2010]. Indeed, study specific differences in comparative physiology are widspread. For example, experimental designs can vary in the temperatures that they apply or in the dosages of chemical modifiers used in experimental treatments [e.g., @WuSeebacher2020]. In addition, they often include a diversity of study species which can vary in responses. To the uninitiated the concern might not be unfounded given that meta-analyses in ecology and evolution often mix highly heterogenous effect size estimates often resulting in high variability [@Senior2016].

@@

## **The Apples and Oranges 'Problem': How lack of comparability can make effects vary widely across studies**
 I'm thinking now, in re-reading this, that we could use a slightly expanded version of paragraph 4 taht describes this problem in a little more detail. This is likely the place for a figure, which I am having some trouble envisaging where one is best placed. It just otherwise seems like a bit of a jump to otherwise head straight into the survey

>>>>

Despite the widespread adoption and well established methodological procedures of meta-analysis it is often criticised for mixing 'apples and oranges' - in otherwords, mixing effects from studies that lack comparability [@Gurevitch2018; @Arnqvist1995; @Gallo1979; @Carpenter2020; @Stewart2010; @Glass2015]. Lack of comparability could result from studies differing in experimental design, species, measurement variables and sampling units, which can impact the generalisability of meta-analytic results [@Arnqvist1995; @Stewart2010]. Indeed, study specific differences in comparative physiology are widspread. For example, experimental designs can vary in the temperatures that they apply or in the dosages of chemical modifiers used in experimental treatments [e.g., @WuSeebacher2020]. In addition, they often include a diversity of study species which can vary in responses. To the uninitiated the concern might not be unfounded given that meta-analyses in ecology and evolution often mix highly heterogenous effect size estimates often resulting in high variability [@Senior2016].

@@

## **The Apples and Oranges 'Problem': How lack of comparability can make effects vary widely across studies**
 I'm thinking now, in re-reading this, that we could use a slightly expanded version of paragraph 4 taht describes this problem in a little more detail. This is likely the place for a figure, which I am having some trouble envisaging where one is best placed. It just otherwise seems like a bit of a jump to otherwise head straight into the survey. Here we should also talk about quality control, the need to be transparent and have careful reporting. This is a big issue in medicine. Talked about a lot. Integrate blurb below...try to take a balanced approach.

In the words of Gene Glass, arguably the father of meta-analysis: "Of course it mixes apples and oranges; in the study of fruit nothing else is sensible; comparing apples and oranges is the only endeavor worthy of true scientists; comparing apples to apples is trivial" [p.g., 224, @Glass2015]. We tend to agree, particularly in comparative physiology. By synthesizing diverse study outcomes, it allows us to not just report on overall effects, but understand critical features of the literature, study systems and approaches that can explain the diversity of effect outcomes observed. In doing so, it provides a richer set of conclusions that can draw attention to important sources of variability that might otherwise be overlooked. Having said that, it is still very improtant to understand the limitations and diversity of studies included within a meta-analysis. Answers to certain questions will inevitably depend on the homogeneity of effects being synthesized.
